Computable Poincar\'e--Friedrichs constants for the $L^{p}$ de Rham complex over convex domains and domains with shellable triangulations

### 2024-09-29 main.tex
shellable triangulations

\documentclass[10pt,letterpaper]{article}
\usepackage[utf8]{inputenc}
\usepackage{fullpage}
%\usepackage[margin=0.1cm]{geometry}
\input{header}

@@

\newcommand{\mwl}[1]{{\color{red} MWL: #1}}
\newcommand{\notice}[1]{{\color{red}REMARK: #1}}
\newcommand{\disk}{\cye{ball}}

@@

\begin{remark}
    Lemma~\ref{lemma:volumecomparison} fully controls the shape quantity $\volumeratio(\calT)$ \cye{of~\eqref{math:volumeratio}} but not $\diameterratio(\calT)$ \cye{of~\eqref{math:diametercomparison}}. 
    Lemma~\ref{lemma:volumecomparison} fully controls the shape quantity $\volumeratio(\calT)$ but not $\diameterratio(\calT)$. 
    As seen above, we can bound the ratio of diameters and volumes of simplices that share at least one face in terms of the shape measure. 
    If we have an upper estimate for the number of simplices sharing a given vertex, 
    we can also bound the ratios of diameters and volumes of simplices sharing only a single common vertex. 
\end{remark}
%%%%%%%% END HERE

@@

\mwl{In this section, we develop a stepwise computable estimate of Poincar\'e--Friedrichs constants of triangulated domains. 
The following very classical procedure serves us as an inspiration: given a gradient vector field, we can reconstruct the scalar potential up to a constant by fixing a starting point and integrating the gradient vector field along lines emanating from that starting point. 
We perform a discrete analogue of this procedure over triangulated domains: 
having fixed a starting triangle, we traverse the triangulation along face-neighboring simplices.
%We construct a scalar potential for a given gradient vector field on larger and larger triangulated subdomains.
We always construct a gradient potential over the new simplex and fix the constant using the value already known on the connecting face, thus constructing a scalar potential over larger and larger subdomains.
This basic idea has appeared in the different variants before, for instance recently in \cite{Brae_Pill_Sch_p_rob_09, ern2020stable, Chaum_Voh_p_rob_3D_H_curl_24, Voh_loc_glob_H1_24}.}

@@

The next auxiliary result extends the results on Poincar\'e--Friedrichs constants over convex domains from Section~\ref{section: PX_convex} to face patches within simplicial triangulations. We point out that face patches are not necessarily convex. The following result is reasonably sharp when the two simplices have similar volumes and diameters.

@@

For the remainder of this section, the following combinatorial notion of triangulations is used. 
Whenever $\calT$ is a face-connected $n$-dimensional triangulation, 
a \emph{spanning tree} is an enumeration $T_{0}, \dots, T_{M}$ of all its $n$-dimensional simplices
where for each index $1 \leq m \leq M$ we have a fixed index $0 \leq \iota(m) \leq m-1$ 
such that $T_{m}$ and $T_{\iota(m)}$ are face-neighboring.
We call $T_{\iota(m)}$ the \emph{precursor} of $T_{m}$ within the spanning tree. 
The reader is referred to Figure~\ref{figure:spanningtree} for an illustration.

@@

\begin{theorem}\label{theorem:poincarefriedrichsestimate:grad}
    Let $\calT$ be a face-connected $n$-dimensional finite triangulation 
    with a spanning tree $T_0, T_1, \dots, T_{M}$. \mwl{should be a sequence with precursor function}
    For any $1 \leq m \leq \cye{M}$, let $F_m = T_m \cap T_{\iota(m)}$.
    Suppose $1 \leq p,q \leq \infty$ with $1 = 1/p + 1/q$,
    and that the domain $\Omega$ is the interior of the underlying set of $\calT$. 
    Then for any $u \in W^{1,p}(\Omega)$ 
    there exists $w \in W^{1,p}(\Omega)$ with $\nabla w = \nabla u$ 
    and such that for all $1 \leq m \leq M$
    we have the following recursive estimates:
    \begin{gather*}
        \| w_0 \|_{L^{p}(T_{0})} \leq C_{{\rm{PF}},T_{0},p} \| \nabla u \|_{L^{p}(T_{0})}.
    \end{gather*}
    and for $1 \leq m \leq M$ we have 
    \begin{align*}
        \| w \|_{L^{p}(T_{m})}
        &
        \leq  
        \volumeratio(\calT)^{\frac 1 p} 
        \| w \|_{L^{p}(T_{\iota(m)})} 
        + 
%         C_{\PF,T_{m},F_{m},p} 
%         \left( 
%             \| \nabla u \|_{L^{p}(T_{m})} 
%             + 
%             \volumeratio(\calT)^{\frac 1 p} \xi(\calT)
%             \| \nabla u \|_{L^{p}(T_{\iota(m)})}
%         \right)
        \left( 1 + \volumeratio(\calT)^{\frac q p} \xi(\calT)^{q} \right)^{\frac 1 q}
        C_{\PF,T_{m},F_{m},p} 
        \| \nabla u \|_{L^{p}(U_{F_m})} 
        ,
        \\
        \| w \|_{L^{p}(T_m)}
        &\leq 
        \volumeratio(\calT)^{\frac 1 p} 
        \| w \|_{L^{p}(T_{{\iota(m)}})}
        +
        \left( 1 + \volumeratio(\calT)^{\frac q p} \right)^{\frac 1 q}
        C_{{\rm{PF}},U_{F_m},p} 
        \| \nabla u \|_{L^{p}(U_{F_m})} 
        . 
    \end{align*}
\end{theorem}

\begin{proof}
    Let $u \in W^{1,p}(\Omega)$. 
    We start with the Poincar\'e--Friedrichs inequality on the first simplex $T_{0}$. 
    There exists $w_0 \in W^{1,p}(T_{0})$ satisfying $\nabla w_0 = \nabla u$ over $T_{0}$ together with 
    \begin{gather*}
    \| w_0 \|_{L^{p}(T_{0})} \leq C_{{\rm{PF}},T_{0},p} \| \nabla u \|_{L^{p}(T_{0})}.
    \end{gather*}
    In particular, $c_{0} := w_0 - u$ is a constant function. 
    We then define 
    \begin{align*}
        w := u + c_0
        .
    \end{align*} 
    Clearly, $w \in W^{1,p}(\Omega)$ with $\nabla w = \nabla u$.
    By construction $w_{|T_{0}} = w_0$. 
    We verify that $w$ satisfies the two desired recursive estimates.

    We continue by discussing the first recursive estimate. 
    Suppose that $1 \leq m \leq M$.
    Recall that we write $F_m := T_m \cap T_{\iota(m)}$, which is a face of dimension $n-1$ shared by the $n$-simplices $T_{m}$ and $T_{m-1}$. 
    We define $w'_{m} := w_{|T_{\iota(m)}} \circ \Xi \in W^{1,p}(T_{m})$,
    where $\Xi : T_{m} \rightarrow T_{\iota(m)}$ is the unique affine diffeomorphism that leaves $F_{m}$ invariant. 
    By construction, $w'_{m} \in W^{1,p}(T_{m})$ with 
    \begin{align*}
        \trace_{F_{m}} w'_{m} = \trace_{F_{m}} w_{|T_{\iota(m)}}
        .
    \end{align*}
    We now define $w''_{m} \in W^{1,p}(T_{m})$ via 
    \begin{align}\label{equation:w''}
        w''_{m} 
        := 
%         u_{|T_{m}} - w'_{m} + c_0 
%         = 
        w_{|T_{m}} - w'_{m} = u_{|T_{m}} - u_{|T_{\iota(m)}} \circ \Xi.
    \end{align}
    We crucially note that $w''_{m}$ is trace-free along $F_{m}$ since
    \begin{align*}
        \trace_{F_{m}} w''_{m} 
        = 
        \trace_{F_{m}} \left( 
            w_{|T_{m}} - w'_{m} 
        \right) 
        =
        \trace_{F_{m}} w_{|T_{m}}
        -
        \trace_{F_{m}} w_{|T_{\iota(m)}}
        =
        \trace_{F_{m}} u_{|T_{m}}
        -
        \trace_{F_{m}} u_{|T_{\iota(m)}}
%         = 
%         \trace_{F_{m}} \left( 
%             u_{|T_{m}} - w'_{m} + c_0 
%         \right) 
%         = 
%         \trace_{F_{m}} \left( 
%             u_{|T_{\iota(m)}} - w_{|T_{\iota(m)}} + c_0 
%         \right) 
        = 0.
    \end{align*}
    We apply Lemma~\ref{lemma:mixedbconsimplex} to the first expression in~\eqref{equation:w''}, which gives 
    \begin{align*}
        \| w''_{m} \|_{L^{p}(T_{m})} 
        &
        \leq 
        C_{\PF,T_{m},F_{m},p} 
        \left( 
            \| \nabla w \|_{L^{p}(T_{m})} 
            + 
            \| \nabla w'_{m} \|_{L^{p}(T_{m})} 
        \right) 
        \leq 
        C_{\PF,T_{m},F_{m},p} 
        \left( 
            \| \nabla u \|_{L^{p}(T_{m})} 
            + 
            \| \nabla w'_{m} \|_{L^{p}(T_{m})} 
        \right) 
        .
    \end{align*}
    Using Lemma~\ref{lemma:volumecomparison} as well as Definitions~\eqref{math:volumeratio} and~\eqref{math:reflectionmeasure}, we find 
    \begin{align*}
        \| \nabla w'_{m} \|_{L^{p}(T_{m})}
        &
        \leq 
        |\det(\Jacobian \Xi  )|^{-\frac 1 p} 
        \matnorm{ \Jacobian \Xi }
        \| \nabla w \|_{L^{p}(T_{\iota(m)})}
        \\&
        \leq 
        \volumeratio(\calT)^{\frac 1 p} \xi(\calT)
        \| \nabla w \|_{L^{p}(T_{\iota(m)})}
        \\&
        =
        \volumeratio(\calT)^{\frac 1 p} \xi(\calT)
        \| \nabla u \|_{L^{p}(T_{\iota(m)})}
        .
    \end{align*}
    % The second expression in~\eqref{equation:w''} shows $w_{|T_{m}} = w''_{m} + w'_{m}$. We finally find 
    Since $w_{|T_{m}} = w''_{m} + w'_{m}$, we finally find 
    \begin{align*}
        \| w \|_{L^{p}(T_{m})}
        &
        \leq  
        \| w'_{m} \|_{L^{p}(T_{m})}
        + 
        \| w''_{m} \|_{L^{p}(T_{m})}
        \\&
        \leq  
        \volumeratio(\calT)^{\frac 1 p} 
        \| w \|_{L^{p}(T_{\iota(m)})} 
        + 
        C_{\PF,T_{m},F_{m},p} 
        \left( 
            \| \nabla u \|_{L^{p}(T_{m})} 
            + 
            \volumeratio(\calT)^{\frac 1 p} \xi(\calT)
        \| \nabla u \|_{L^{p}(T_{\iota(m)})}
        \right). 
    \end{align*}
    We now use H\"older's inequality. 
    Therefrom, the first recursive estimate follows.

    Now we discuss the second recursive estimate. 
    Suppose again that $1 \leq m \leq M$. 
    Recall that we write $U_{F_m} := T_m \cup T_{\iota(m)}$. 
    We use the Poincar\'e--Friedrichs inequality over $U_{F_{m}}$,
    as given in Lemma~\ref{lemma:poincarefriedrichsoverfacepatch}, 
    to find $w_{F_m} \in W^{1,p}(U_{F_m})$ such that $\nabla w_{F_m} = \nabla u$ over $U_{F_m}$ and 
    \begin{align}\label{math:wFm_inequality}
        \| w_{F_m} \|_{L^{p}(U_{F_m})} \leq C_{{\rm{PF}},U_{F_m},p} \| \nabla w_{F_m} \|_{L^{p}(U_{F_m})}.
    \end{align}
%     \cye{Since $\nabla (w|_{U_{F_m}}) = \nabla u_{F_{m}} = \nabla (u|_{U_{F_m}})$, 
    We can define the constant
    \begin{align*}
    c_{m} := w|_{U_{F_m}} - w_{F_{m}}.
    \end{align*}
    We now observe that 
    \begin{gather*}
        \| w \|_{L^{p}(T_m)}
        \leq 
        \| w_{F_m} \|_{L^{p}(T_m)}
        +
        \| c_{m} \|_{L^{p}(T_m)},
        \\
        \| c_{m} \|_{L^{p}(T_m)}
        = 
        \frac{ \vol(T_m)^{\frac 1 p} }{ \vol(T_{{\iota(m)}})^{\frac 1 p} }
        \| c_{m} \|_{L^{p}(T_{{\iota(m)}})},
        \\ 
        \| c_{m} \|_{L^{p}(T_{{\iota(m)}})}
        \leq 
        \| w \|_{L^{p}(T_{{\iota(m)}})} + \| w_{F_{m}} \|_{L^{p}(T_{{\iota(m)}})} 
        .
    \end{gather*}
    In combination, 
    \begin{align*}
        \| w \|_{L^{p}(T_m)}
        &
        \leq 
        \| w_{F_m} \|_{L^{p}(T_m)}
        \\&\qquad 
        +
        \frac{ \vol(T_m)^{\frac 1 p} }{ \vol(T_{{\iota(m)}})^{\frac 1 p} }
        \| w_{F_{m}} \|_{L^{p}(T_{{\iota(m)}})}
        +
        \frac{ \vol(T_m)^{\frac 1 p} }{ \vol(T_{{\iota(m)}})^{\frac 1 p} }
        \| w \|_{L^{p}(T_{{\iota(m)}})}
        .
    \end{align*}
    We sum the two integrals of $w_{F_m}$. 
    % 
    % 1/q = 1 - 1/p = (p-1)/p 
    When $1 < p < \infty$, recalling the complementary exponent $q = p/(p-1) \in (1,\infty)$, 
    we use H\"older's inequality to verify 
    \begin{align*}
        \| w \|_{L^{p}(T_m)}
        &\leq 
        \left( 1 + \frac{ \vol(T_m)^{\frac q p} }{ \vol(T_{{\iota(m)}})^{\frac q p} } \right)^{\frac 1 q}
        \left( 
            \| w_{F_{m}} \|_{L^{p}(T_m)}^{p}
            +
            \| w_{F_{m}} \|_{L^{p}(T_{{\iota(m)}})}^{p}
        \right)^{\frac 1 p}
        +
        \frac{ \vol(T_m)^{\frac 1 p} }{ \vol(T_{{\iota(m)}})^{\frac 1 p} }
        \| w \|_{L^{p}(T_{{\iota(m)}})}
        \\&\leq 
        \left( 1 + \frac{ \vol(T_m)^{\frac q p} }{ \vol(T_{{\iota(m)}})^{\frac q p} } \right)^{\frac 1 q}
        \| w_{F_{m}} \|_{L^{p}(U_{F_m})} 
        +
        \frac{ \vol(T_m)^{\frac 1 p} }{ \vol(T_{{\iota(m)}})^{\frac 1 p} }
        \| w \|_{L^{p}(T_{{\iota(m)}})}
        .
    \end{align*}
    In the special case $p=1$,
    \begin{align*}
        \| w\|_{L^{1}(T_m)}
        \leq 
        \max\left(
            1, \frac{ \vol(T_m) }{ \vol(T_{{\iota(m)}}) } 
        \right)
        \| w_{F_m} \|_{L^{1}(U_{F_m})}
        +
        \frac{ \vol(T_m) }{ \vol(T_{{\iota(m)}}) }
        \| w \|_{L^{1}(T_{{\iota(m)}})}
        .
    \end{align*}
    In the limit case $p=\infty$, 
    \begin{align*}
        \| w \|_{L^{\infty}(T_m)}
        \leq 
        2
        \| w_{F_m} \|_{L^{\infty}(U_{F_m})}
        +
        \| w \|_{L^{\infty}(T_{{\iota(m)}})}
        .
    \end{align*}
    The local inequality~\eqref{math:wFm_inequality} now provides the second recursive estimate. 
    The proof is complete.
\end{proof}

@@

\begin{remark}
    The Poincar\'e--Friedrichs constant obtained from Theorem~\ref{theorem:poincarefriedrichsestimate:grad} increasingly overestimates the best one as number of $n$-simplices in the triangulation $\calT$ increases. We hence target its use to local patches (stars), in particular nonconvex boundary stars. The latter occur inevitably at reentrant corners. Clearly, the same building principle applies whenever we have any nonoverlapping partition of $\overline \Omega$ into convex local patches of $n$-simplices $U_m$ with a face-connected path. The proof proceeds verbatim, where we merely replace the simplices $T_{m}$ by the convex local patches $U_m$. This may allow to a large decrease of the number of elements in the partition of $\overline \Omega$ and thus to a largely improved estimate of the best Poincar\'e--Friedrichs constant. 
\end{remark}

@@

We return to the theory of triangulations,
as our main objective requires some further concepts.
We are interested in simplicial complexes that triangulate manifolds. 
Our particular interest are manifold triangulations that are \emph{shellable}.
Such simplicial complexes are constructed by successively adding simplices in a particularly well-behaved manner. 
\cye{Local} patches \cye{(stars)} within triangulations of dimension two and three are examples of such shellable complexes. 
The monographs by Kozlov~\cite{kozlov2008combinatorial} and Ziegler~\cite{ziegler1995lectures} are our main references for this section. 
We also refer to Lee's monograph~\cite{lee2011topological} for any further background on manifolds.

\begin{figure}[h]
\centerline{\includegraphics[width=0.2\textwidth]{revision/homotopic_ball.jpg} \quad \includegraphics[width=0.2\textwidth]{revision/homotopic_ball.jpg}}
\label{figure:not_manifold_triang}
\caption{A manifold triangulation (left) and not a manifold triangulation (right)}
\end{figure}

@@

    \item
    Let $F \in \Faces(T)$ and let $z_F \in F$ be its barycenter. 
    Since $\calT$ is finite, we let $\mathring B_F$ be an open neighborhood around $z_F$ so small that it only intersects those $n$-simplices of $\calT$ that already contain $z_F$ and no faces other than $F$.
    Suppose there are distinct $n$-simplices $T_1, T_2, \dots, T_K$ that contain $z_F$. 
    The intersection of any two of them is $F$, but their interiors are disjoint because otherwise they would coincide.

@@

The reason of our interest in shellable simplicial complexes is that they can be constructed by \cye{a} successive adhesion of simplices to intermediate complexes \cye{where these intermediate complexes are} simplicial \disk s \cye{\sout{or a simplicial spheres}}, as shows the following important consequence.\todo{MV: OK? ``Intermediate'' excludes the last one.}

>>>>

\documentclass[10pt,letterpaper]{article}
\usepackage[utf8]{inputenc}
\usepackage{fullpage}
\usepackage[margin=0.1cm]{geometry}
\input{header}

@@

\newcommand{\mwl}[1]{{\color{red}#1}}
\newcommand{\notice}[1]{{\color{red}REMARK: #1}}
\newcommand{\disk}{\cye{ball}}

@@

\begin{remark}
    Lemma~\ref{lemma:volumecomparison} fully controls the shape quantity $\volumeratio(\calT)$ \cye{of~\eqref{math:volumeratio}} but not $\diameterratio(\calT)$ \cye{of~\eqref{math:diameterratio}}. 
    Lemma~\ref{lemma:volumecomparison} fully controls the shape quantity $\volumeratio(\calT)$ but not $\diameterratio(\calT)$. 
    As seen above, we can bound the ratio of diameters and volumes of simplices that share at least one face in terms of the shape measure. 
    If we have an upper estimate for the number of simplices sharing a given vertex, 
    we can also bound the ratios of diameters and volumes of simplices sharing only a single common vertex. 
\end{remark}
%%%%%%%% END HERE

@@

In this section, we develop a stepwise computable estimate of Poincar\'e--Friedrichs constants of triangulated domains. 
The following very classical procedure serves us as an inspiration: given a gradient vector field, we can reconstruct the scalar potential up to a constant by fixing a starting point and integrating the gradient vector field along lines emanating from that starting point. 
We perform a discrete analogue of this procedure over triangulated domains: 
having fixed a starting triangle, we traverse the triangulation along face-neighboring simplices.
%We construct a scalar potential for a given gradient vector field on larger and larger triangulated subdomains.
We always construct a gradient potential over the new simplex and fix the constant using the value already known on the connecting face, thus constructing a scalar potential over larger and larger subdomains.
This basic idea has appeared in the different variants before, for instance recently in \cite{Brae_Pill_Sch_p_rob_09, ern2020stable, Chaum_Voh_p_rob_3D_H_curl_24, Voh_loc_glob_H1_24}.

@@

The next auxiliary result extends the results on Poincar\'e--Friedrichs constants over convex domains from Section~\ref{section: PX_convex} to face patches within simplicial triangulations. We emphasize that face patches are not necessarily convex. The following result is reasonably sharp when the two simplices have similar volumes and diameters.

@@

\begin{comment}
For the remainder of this section, the following combinatorial notion of triangulations is used. 
Whenever $\calT$ is a face-connected $n$-dimensional triangulation, 
a \emph{spanning tree} is an enumeration $T_{0}, \dots, T_{M}$ of all its $n$-dimensional simplices
where for each index $1 \leq m \leq M$ we have a fixed index $0 \leq \iota(m) \leq m-1$ 
such that $T_{m}$ and $T_{\iota(m)}$ are face-neighboring.
We call $T_{\iota(m)}$ the \emph{precursor} of $T_{m}$ within the spanning tree. 
The reader is referred to Figure~\ref{figure:spanningtree} for an illustration. 
\end{comment}

@@

\begin{theorem}\label{theorem:poincarefriedrichsestimate:grad}
    Let $\calT$ be a face-connected $n$-dimensional finite triangulation. 
    Suppose $1 \leq p,q \leq \infty$ with $1 = 1/p + 1/q$,
    and that the domain $\Omega$ is the interior of the underlying set of $\calT$. 
    Then for any $u \in W^{1,p}(\Omega)$ 
    there exists $w \in W^{1,p}(\Omega)$ with $\nabla w = \nabla u$ 
    and satisfying the following estimates:
    there exists an $n$-simplex $T_0 \in \calT$ with 
    \begin{gather*}
        \| w_0 \|_{L^{p}(T_{0})} \leq C_{{\rm{PF}},T_{0},p} \| \nabla u \|_{L^{p}(T_{0})}.
    \end{gather*}
    and whenever $T_0, T_1, \dots, T_M$ is a face path,
    then for all $1 \leq m \leq$ we the following recursive estimates:
    \begin{align*}
        \| w \|_{L^{p}(T_{m})}
        &
        \leq  
        \volumeratio(\calT)^{\frac 1 p} 
        \| w \|_{L^{p}(T_{\iota(m)})} 
        + 
%         C_{\PF,T_{m},F_{m},p} 
%         \left( 
%             \| \nabla u \|_{L^{p}(T_{m})} 
%             + 
%             \volumeratio(\calT)^{\frac 1 p} \xi(\calT)
%             \| \nabla u \|_{L^{p}(T_{\iota(m)})}
%         \right)
        \left( 1 + \volumeratio(\calT)^{\frac q p} \xi(\calT)^{q} \right)^{\frac 1 q}
        C_{\PF,T_{m},F_{m},p} 
        \| \nabla u \|_{L^{p}(U_{F_m})} 
        ,
        \\
        \| w \|_{L^{p}(T_m)}
        &\leq 
        \volumeratio(\calT)^{\frac 1 p} 
        \| w \|_{L^{p}(T_{{\iota(m)}})}
        +
        \left( 1 + \volumeratio(\calT)^{\frac q p} \right)^{\frac 1 q}
        C_{{\rm{PF}},U_{F_m},p} 
        \| \nabla u \|_{L^{p}(U_{F_m})} 
        . 
    \end{align*}
    Here, for any $1 \leq m \leq M$, let $F_m = T_m \cap T_{m-1}$.
\end{theorem}

\begin{proof}
    Let $u \in W^{1,p}(\Omega)$. 
    We start with the Poincar\'e--Friedrichs inequality on the first simplex $T_{0}$. 
    There exists $w_0 \in W^{1,p}(T_{0})$ satisfying $\nabla w_0 = \nabla u$ over $T_{0}$ together with 
    \begin{gather*}
        \| w_0 \|_{L^{p}(T_{0})} \leq C_{{\rm{PF}},T_{0},p} \| \nabla u \|_{L^{p}(T_{0})}.
    \end{gather*}
    In particular, $c_{0} := w_0 - u$ is a constant function. 
    We then define 
    \begin{align*}
        w := u + c_0
        .
    \end{align*} 
    Clearly, $w \in W^{1,p}(\Omega)$ with $\nabla w = \nabla u$.
    By construction $w_{|T_{0}} = w_0$. 
    We verify that $w$ satisfies the two desired recursive estimates.
    Suppose that $T_0, T_1, \dots, T_M$ is a face path in $\calT$
    and that $1 \leq m \leq M$.
    Recall that we write $F_m := T_m \cap T_{m-1}$, which is a face of dimension $n-1$ shared by the $n$-simplices $T_{m}$ and $T_{m-1}$.

    We proceed with showing the first recursive estimate. 
    We define $w'_{m} := w_{|T_{m-1}} \circ \Xi \in W^{1,p}(T_{m})$,
    where $\Xi : T_{m} \rightarrow T_{m-1}$ is the unique affine diffeomorphism that leaves $F_{m}$ invariant. 
    By construction, $w'_{m} \in W^{1,p}(T_{m})$ with 
    \begin{align*}
        \trace_{F_{m}} w'_{m} = \trace_{F_{m}} w_{|T_{m-1}}
        .
    \end{align*}
    We now define $w''_{m} \in W^{1,p}(T_{m})$ via 
    \begin{align}\label{equation:w''}
        w''_{m} 
        := 
%         u_{|T_{m}} - w'_{m} + c_0 
%         = 
        w_{|T_{m}} - w'_{m} = u_{|T_{m}} - u_{|T_{m-1}} \circ \Xi.
    \end{align}
    We crucially note that $w''_{m}$ is trace-free along $F_{m}$ since
    \begin{align*}
        \trace_{F_{m}} w''_{m} 
        = 
        \trace_{F_{m}} \left( 
            w_{|T_{m}} - w'_{m} 
        \right) 
        =
        \trace_{F_{m}} w_{|T_{m}}
        -
        \trace_{F_{m}} w_{|T_{m-1}}
        =
        \trace_{F_{m}} u_{|T_{m}}
        -
        \trace_{F_{m}} u_{|T_{m-1}}
%         = 
%         \trace_{F_{m}} \left( 
%             u_{|T_{m}} - w'_{m} + c_0 
%         \right) 
%         = 
%         \trace_{F_{m}} \left( 
%             u_{|T_{m-1}} - w_{|T_{m-1}} + c_0 
%         \right) 
        = 0.
    \end{align*}
    We apply Lemma~\ref{lemma:mixedbconsimplex} to the first expression in~\eqref{equation:w''}, which gives 
    \begin{align*}
        \| w''_{m} \|_{L^{p}(T_{m})} 
        &
        \leq 
        C_{\PF,T_{m},F_{m},p} 
        \left( 
            \| \nabla w \|_{L^{p}(T_{m})} 
            + 
            \| \nabla w'_{m} \|_{L^{p}(T_{m})} 
        \right) 
        \leq 
        C_{\PF,T_{m},F_{m},p} 
        \left( 
            \| \nabla u \|_{L^{p}(T_{m})} 
            + 
            \| \nabla w'_{m} \|_{L^{p}(T_{m})} 
        \right) 
        .
    \end{align*}
    Using Lemma~\ref{lemma:volumecomparison} as well as Definitions~\eqref{math:volumeratio} and~\eqref{math:reflectionmeasure}, we find 
    \begin{align*}
        \| \nabla w'_{m} \|_{L^{p}(T_{m})}
        &
        \leq 
        |\det(\Jacobian \Xi  )|^{-\frac 1 p} 
        \matnorm{ \Jacobian \Xi }
        \| \nabla w \|_{L^{p}(T_{m-1})}
        \\&
        \leq 
        \volumeratio(\calT)^{\frac 1 p} \xi(\calT)
        \| \nabla w \|_{L^{p}(T_{m-1})}
        \\&
        =
        \volumeratio(\calT)^{\frac 1 p} \xi(\calT)
        \| \nabla u \|_{L^{p}(T_{m-1})}
        .
    \end{align*}
    % The second expression in~\eqref{equation:w''} shows $w_{|T_{m}} = w''_{m} + w'_{m}$. We finally find 
    Since $w_{|T_{m}} = w''_{m} + w'_{m}$, we finally find 
    \begin{align*}
        \| w \|_{L^{p}(T_{m})}
        &
        \leq  
        \| w'_{m} \|_{L^{p}(T_{m})}
        + 
        \| w''_{m} \|_{L^{p}(T_{m})}
        \\&
        \leq  
        \volumeratio(\calT)^{\frac 1 p} 
        \| w \|_{L^{p}(T_{m-1})} 
        + 
        C_{\PF,T_{m},F_{m},p} 
        \left( 
            \| \nabla u \|_{L^{p}(T_{m})} 
            + 
            \volumeratio(\calT)^{\frac 1 p} \xi(\calT)
        \| \nabla u \|_{L^{p}(T_{m-1})}
        \right). 
    \end{align*}
    We now use H\"older's inequality. 
    Therefrom, the first recursive estimate follows.

    Now we discuss the second recursive estimate. 
    Suppose again that $1 \leq m \leq M$. 
    Recall that we write $U_{F_m} := T_m \cup T_{m-1}$. 
    We use the Poincar\'e--Friedrichs inequality over $U_{F_{m}}$,
    as given in Lemma~\ref{lemma:poincarefriedrichsoverfacepatch}, 
    to find $w_{F_m} \in W^{1,p}(U_{F_m})$ such that $\nabla w_{F_m} = \nabla u$ over $U_{F_m}$ and 
    \begin{align}\label{math:wFm_inequality}
        \| w_{F_m} \|_{L^{p}(U_{F_m})} \leq C_{{\rm{PF}},U_{F_m},p} \| \nabla w_{F_m} \|_{L^{p}(U_{F_m})}.
    \end{align}
%     \cye{Since $\nabla (w|_{U_{F_m}}) = \nabla u_{F_{m}} = \nabla (u|_{U_{F_m}})$, 
    We can define the constant
    \begin{align*}
    c_{m} := w|_{U_{F_m}} - w_{F_{m}}.
    \end{align*}
    We now observe that 
    \begin{gather*}
        \| w \|_{L^{p}(T_m)}
        \leq 
        \| w_{F_m} \|_{L^{p}(T_m)}
        +
        \| c_{m} \|_{L^{p}(T_m)},
        \\
        \| c_{m} \|_{L^{p}(T_m)}
        = 
        \frac{ \vol(T_m)^{\frac 1 p} }{ \vol(T_{{m-1}})^{\frac 1 p} }
        \| c_{m} \|_{L^{p}(T_{{m-1}})},
        \\ 
        \| c_{m} \|_{L^{p}(T_{{m-1}})}
        \leq 
        \| w \|_{L^{p}(T_{{m-1}})} + \| w_{F_{m}} \|_{L^{p}(T_{{m-1}})} 
        .
    \end{gather*}
    In combination, 
    \begin{align*}
        \| w \|_{L^{p}(T_m)}
        &
        \leq 
        \| w_{F_m} \|_{L^{p}(T_m)}
        \\&\qquad 
        +
        \frac{ \vol(T_m)^{\frac 1 p} }{ \vol(T_{{m-1}})^{\frac 1 p} }
        \| w_{F_{m}} \|_{L^{p}(T_{{m-1}})}
        +
        \frac{ \vol(T_m)^{\frac 1 p} }{ \vol(T_{{m-1}})^{\frac 1 p} }
        \| w \|_{L^{p}(T_{{m-1}})}
        .
    \end{align*}
    We sum the two integrals of $w_{F_m}$. 
    % 
    % 1/q = 1 - 1/p = (p-1)/p 
    When $1 < p < \infty$, recalling the complementary exponent $q = p/(p-1) \in (1,\infty)$, 
    we use H\"older's inequality to verify 
    \begin{align*}
        \| w \|_{L^{p}(T_m)}
        &\leq 
        \left( 1 + \frac{ \vol(T_m)^{\frac q p} }{ \vol(T_{{m-1}})^{\frac q p} } \right)^{\frac 1 q}
        \left( 
            \| w_{F_{m}} \|_{L^{p}(T_m)}^{p}
            +
            \| w_{F_{m}} \|_{L^{p}(T_{{m-1}})}^{p}
        \right)^{\frac 1 p}
        +
        \frac{ \vol(T_m)^{\frac 1 p} }{ \vol(T_{{m-1}})^{\frac 1 p} }
        \| w \|_{L^{p}(T_{{m-1}})}
        \\&\leq 
        \left( 1 + \frac{ \vol(T_m)^{\frac q p} }{ \vol(T_{{m-1}})^{\frac q p} } \right)^{\frac 1 q}
        \| w_{F_{m}} \|_{L^{p}(U_{F_m})} 
        +
        \frac{ \vol(T_m)^{\frac 1 p} }{ \vol(T_{{m-1}})^{\frac 1 p} }
        \| w \|_{L^{p}(T_{{m-1}})}
        .
    \end{align*}
    In the special case $p=1$,
    \begin{align*}
        \| w\|_{L^{1}(T_m)}
        \leq 
        \max\left(
            1, \frac{ \vol(T_m) }{ \vol(T_{{m-1}}) } 
        \right)
        \| w_{F_m} \|_{L^{1}(U_{F_m})}
        +
        \frac{ \vol(T_m) }{ \vol(T_{{m-1}}) }
        \| w \|_{L^{1}(T_{{m-1}})}
        .
    \end{align*}
    In the limit case $p=\infty$, 
    \begin{align*}
        \| w \|_{L^{\infty}(T_m)}
        \leq 
        2
        \| w_{F_m} \|_{L^{\infty}(U_{F_m})}
        +
        \| w \|_{L^{\infty}(T_{{m-1}})}
        .
    \end{align*}
    The local inequality~\eqref{math:wFm_inequality} now provides the second recursive estimate. 
    The proof is complete.
\end{proof}

@@

\begin{remark}
    The Poincar\'e--Friedrichs constant obtained from Theorem~\ref{theorem:poincarefriedrichsestimate:grad} increasingly overestimates the best one as number of $n$-simplices in the triangulation $\calT$ increases. We hence target its use to local patches (stars), in particular nonconvex boundary stars. The latter occur inevitably at reentrant corners. Clearly, the same building principle applies whenever we have any nonoverlapping partition of $\overline \Omega$ into convex local patches of $n$-simplices $\{ U_m \}$ with a face-connected path. The proof proceeds verbatim, where we merely replace the simplices $\{ T_{m} \}$ by the convex local patches $\{ U_m \}$. This may allow to a large decrease of the number of elements in the partition of $\overline \Omega$ and thus to a largely improved estimate of the best Poincar\'e--Friedrichs constant. 
\end{remark}

@@

We return to the theory of triangulations,
as our main objective requires some further concepts.
We are interested in simplicial complexes that triangulate manifolds. 
Our particular interest are manifold triangulations that are \emph{shellable}.
Such simplicial complexes are constructed by successively adding simplices in a particularly well-behaved manner. 
Local patches (stars) within triangulations of dimension two and three are examples of such shellable complexes. 
The monographs by Kozlov~\cite{kozlov2008combinatorial} and Ziegler~\cite{ziegler1995lectures} are our main references for this section. 
We also refer to Lee's monograph~\cite{lee2011topological} for any further background on manifolds.

\begin{figure}[h]
\centerline{\includegraphics[width=0.2\textwidth]{revision/homotopic_ball.jpg} \quad \includegraphics[width=0.2\textwidth]{revision/homotopic_ball.jpg}}
\caption{A manifold triangulation (left) and not a manifold triangulation (right)}
\label{figure:not_manifold_triang}
\end{figure}

@@

    \item
    Let $F \in \Faces(T)$ and let $z_F \in F$ be its barycenter. 
    Since $\calT$ is finite, we let $\mathring B_F$ be an open neighborhood around $z_F$ 
    homeomorphic to an $n$-dimensional ball 
    so small that its closure $\overline{B_F}$ only intersects those $n$-simplices of $\calT$ that already contain $z_F$ and no faces other than $F$.
    Suppose there are distinct $n$-simplices $T_1, T_2, \dots, T_K$ that contain $z_F$. 
    The intersection of any two of them is $F$, but their interiors are disjoint because otherwise they would coincide.

@@

The reason of our interest in shellable simplicial complexes is that they can be constructed via successive adhesion of simplices.
The resulting succession of simplicial complexes consists of simplicial balls or spheres.
